# SEM HW5 - code in the chunk but don't run them. copy them as a whole with the comments
```{r}
install.packages("lavaan")
install.packages("semPlot")
install.packages('haven')
```


```{r}

library(haven)
library(lavaan)
library(semPlot)
```


```{r}
data <- read_sav('E:/Cloud/OneDrive - University of Missouri/Mizzou_PhD/Class plan/2022 Measurement Seminar Singh/DBAP643_Spring_2021 Sohana_Observed_Data_w_SDRS.sav')
```


## Replicate HW3


```{r}
#--CFA Models

# the syntax is special. It has the '' on top 
#one-factor model with loading fixed to 1
model1<-'RD=~rd1 + rd2 + rd3 + rd4'
fit1<-cfa(model1,data=data)
summary(fit1,fit.measures=TRUE)
parameterEstimates(fit1,standardized=TRUE)
semPaths(fit1,what="paths",whatLabels="par")
```

```{r}
library(dplyr) 
# Standardize the variables 
#data_std <- data |> mutate_at(c('rd1', 'rd2'), ~(scale(.) |> as.vector))
#data_std
data_std <- data %>% mutate_at(c('rd1', 'rd2','rd3','rd4','ra1','ra2','rc1','rc2','rc3',
'cr1','cr2','cr3','cr4','lc1','lc2','lc3','lc4','de1','de2',
'ap1','ap2','ap3','SDRS1','SDRS2','SDRS3','SDRS4','SDRS5'), ~(scale(.) %>% as.vector))


#pip operator, what's the ~ before scale? what's difference between %>% and |>, not the same?
#https://stackoverflow.com/questions/67744604/what-does-pipe-greater-than-mean-in-r
#data_std <- data |> mutate_at(c('rd1', 'rd2'), ~(scale(.) |> as.vector))
#data_std
```


Stata code: 
sem (RD -> rd1std - rd4std) (RA -> ra1std ra2std) (RC -> rc1std - rc3std) , latent( RD  RA  RC )
estat gof, stats(all)

```{r}

#--Three factors CFA Models

model1<-'
RD=~rd1 + rd2 + rd3 + rd4
RA=~ ra1 + ra2
RC=~ rc1 + rc2 + rc3
'
fit1<-cfa(model1,data=data_std)
summary(fit1,fit.measures=TRUE)
parameterEstimates(fit1,standardized=TRUE)
semPaths(fit1,what="paths",whatLabels="par")

```


## most of the previous steps can be done in R and stata in parellel. 
lavaan. 
https://stats.oarc.ucla.edu/r/seminars/rcfa/#s2c


```{r}
library(installr)
```


## output the write up in the R. Only add Stata as complement. -> use TC as template. 


## Goal is to understand the application of regression at different angle to fully understand it.
```{r}
library(lavaan)
```

```{r}
out <- lavaan::mplus2lavaan('E:/Cloud/OneDrive - University of Missouri/Mizzou_PhD/Class plan/2022 Measurement Seminar Singh/HW5/sample0.inp')
#summary(out)
```
```{r}

```

